---
title: "IC Detection Limit - Function Integrated"
format: html
editor: visual
---

## Purpose

This code calculates detection limits for the Metrohm IC in the Bergstrom Biogeochem Lab using the following equation:

$$Detection Limit = 3.3*sigma/m$$

where sigma is the standard deviation of the residuals (the difference in observed results from linear model)

and $$m$$=slope of calibration curve

## Load Data

1.  Update the working directory to the folder that your export data lives in.
2.  Ensure that your correct calibration export .csv files are associated with "cat" and "an"
3.  Upload your run files as cat. and an.samples

```{r}
wd <-setwd("/Users/hannahrichardson/Documents/BSU/Research/Data/Raw Data/IC data")

cat <- read.csv("CATION_CALIBRATION_18Aug_Area.csv")
an <- read.csv("ANION_CALIBRATION_18Aug_Area.csv")
cat.samples<- read.csv('WOLV_Cations_18AUG22.csv')
an.samples <- read.csv('WOLV_AN_18AUG22.csv')
```

## Use DL function

All results are in mg/L

```{r}
IC_detectionlimit(cat,an)
```
