---
title: "IC_Error_Final"
format: html
editor: visual
use: Process data from Metrohm Ion Chromatograph to return detection limits and standard deviations
---

Code developed by Bryn Chaffee adapted from Hannah Richardson

Last edited by: Isabella Welk 28Oct25

**Detection limit equation:**

$$Detection Limit = 3.3*sigma/m$$

| sigma = standard deviation of the residuals
| m = slope of calibration curve
| residuals = difference in observed results from linear model

Error is calculated as the mean difference between residuals (duplicates)

Immediately upon running this code, processed output data files should be uploaded to appropriate project folder (a separate location than the raw data files).

## 1. Set up working environment

```{r}
##clear workspace 
rm(list = ls())

##set working directory -- Where your copy of the code is stored locally
setwd("/Users/isabellawelk/Documents/GitHub/Detection-Limits copy") 
list.files()

#load libraries 
library(dplyr) 
library(reshape) 
library(readr) 
library(googledrive) # Make sure googledrive package is installed, if not, run the following line one time 
#install.packages("googledrive") 
```

## 2. Import .csv Data Files

Exported sample data SHOULD NOT include Areas. Exported Calibration Data SHOULD include Areas.

All data files should be stored in the Bergstrom Watershed Biogeochemistry Lab shared Google Drive. Data can be imported:

1.  Directly from Google Drive (preferred)

2.  From locally stored files

**Define your file names** - These should exactly match what the raw data file is called, and will then be used to name the final exported file.

```{r}
anion_file <- "Wapiti_13Oct25_Anions.csv" 
cation_file <- "Wapiti_13Oct25_Cations.csv"

an_cal_file <- "An_Calibration_13Oct25.csv" 
cat_cal_file <- "Cat_Calibration_13Oct25.csv"
```

#### Upload from Google Drive

This first cell allows you to specify which project folder the data is stored in. The calibrations files should be stored in Standards_Runs \> Anion/Cation folders. The sample data files should be stored in a project folder.

```{r}
project_folder <- "Wapiti" # Change this to match the project folder where the data is located
```

This second cell allows you to pull the files into the Environment without downloading locally

```{r}
# This may prompt you to login on google drive in the console
source("read_drive_csv.R") 

# Samples from project folder
anions <- read_drive_csv(project_path$id, anion_file, sep = ";")
cations <- read_drive_csv(project_path$id, cation_file, sep = ";")

# Calibrations from Standards_Runs
an_calibration <- read_drive_csv(anion_cal_folder$id, an_cal_file, sep = ";")
cat_calibration <- read_drive_csv(cation_cal_folder$id, cat_cal_file, sep = ";")
```

#### Upload Files from Local Storage

Change the file paths to match where the files are stored in your machine.

```{r}
# Or read data from locally stored location
anions <- read_csv("ICRunData/Samples/Starkey_8Aug25_Anions.csv", sep = ';') #anion file
cations <- read_csv("ICRunData/Samples/Starkey_5Aug25_Cations.csv", sep = ';') #cation file

an_calibration <- read_csv("ICRunData/Calibrations/An_Calibration_7Aug25.csv", sep = ';')     #import calibration used for current anion samples
cat_calibration <- read_csv("ICRunData/Calibrations/Cat_Calibration_16July25.csv", sep = ';')  #import calibration used for current cation samples

```

## 3. Calculate detection limit and error

The final files will be exported to a folder called Processed_IC_Data located in the working directory. Files will be named after the `anion_file` and `cation_file` defined in step 2 with '\_processed' added.

If you would like to export files to a different location, you can manually define the file path and run the third cell below.

To calculate the detection limits and standard deviation, run either the cell labeled **files with 'Ident' column** or **files without 'Ident' column** depending on if your data/calibrations has the Ident column exported. (files exported after 15Oct25 should have Ident exported)

#### Files with 'Ident' column (exported post 15Oct25)

```{r}
#specify column names
an.names <- c("Fluoride","Chloride","Nitrite", "Bromide","Nitrate","Phosphate","Sulfate")
cat.names <- c("Lithium","Sodium","Ammonium","Potassium","Magnesium","Calcium")

##clean the data to remove standards, blanks, and invalid values
an.clean <- anions[which(anions$Sample.type == "Sample"),]    #checks the sample.type column and pulls only "Sample" values
cat.clean <- cations[which(cations$Sample.type == "Sample"),]

an.clean <- replace(an.clean, an.clean == 'invalid', NA)   #replaces any 'invalid' values with NA
cat.clean <- replace(cat.clean, cat.clean == 'invalid', NA)

##Doing the same cleaning but for the calibration run
an_calibration  <- an_calibration  %>% filter(grepl("Standard", `Sample.type`, ignore.case = TRUE))
cat_calibration <- cat_calibration %>% filter(grepl("Standard", `Sample.type`, ignore.case = TRUE))

an_calibration <- an_calibration[an_calibration$`Sample.type` != "Check standard 1", ]
cat_calibration <- cat_calibration[cat_calibration$`Sample.type` != "Check standard 1", ]

an_calibration <- replace(an_calibration, an_calibration == 'invalid', NA)   #replaces any 'invalid' values with NA
cat_calibration <- replace(cat_calibration, cat_calibration == 'invalid', NA)


#remove '-duplicate' from any sample names
an.clean$Ident <- gsub("-duplicate", "", as.character(an.clean$Ident))
an.clean$Ident <- gsub("-Duplicate", "", as.character(an.clean$Ident))
an.clean$Ident <- gsub("-DUPLICATE", "", as.character(an.clean$Ident))
an.clean$Ident <- gsub(" DUPLICATE", "", as.character(an.clean$Ident))
an.clean$Ident <- gsub(" Duplicate", "", as.character(an.clean$Ident))
an.clean$Ident <- gsub(" duplicate", "", as.character(an.clean$Ident))
an.clean$Ident <- gsub("_Duplicate", "", as.character(an.clean$Ident))
an.clean$Ident <- gsub("_duplicate", "", as.character(an.clean$Ident))
colnames(an.clean) <- c("Determination.start","Sample.type","Ident","Info.1","Fluoride Concentration","Chloride Concentration","Nitrite Concentration", "Bromide Concentration",
                       "Nitrate Concentration","Phosphate Concentration","Sulfate Concentration")

cat.clean$Ident <- gsub("-duplicate", "", as.character(cat.clean$Ident))
cat.clean$Ident <- gsub("-Duplicate", "", as.character(cat.clean$Ident))
cat.clean$Ident <- gsub("-DUPLICATE", "", as.character(cat.clean$Ident))
cat.clean$Ident <- gsub(" duplicate", "", as.character(cat.clean$Ident))
cat.clean$Ident <- gsub(" Duplicate", "", as.character(cat.clean$Ident))
cat.clean$Ident <- gsub(" DUPLICATE", "", as.character(cat.clean$Ident))
cat.clean$Ident <- gsub("_Duplicate", "", as.character(cat.clean$Ident))
cat.clean$Ident <- gsub("_duplicate", "", as.character(cat.clean$Ident))
cat.clean$Ident <- gsub(" Deuplicate", "", as.character(cat.clean$Ident))
colnames(cat.clean) <- c("Determination.start","Sample.type","Ident","Info.1","Lithium Concentration","Sodium Concentration","Ammonium Concentration",
                        "Potassium Concentration","Magnesium Concentration","Calcium Concentration")

#create cation and anion concentration variables
an.conc <- subset(an.clean, select = -Sample.type)
cat.conc <- subset(cat.clean, select = -Sample.type)



#standard deviation workflow:

#find duplicates
an.dups <- an.clean[duplicated(an.clean$Ident) | duplicated(an.clean$Ident, fromLast = TRUE),]  #locates any duplicated samples within an.clean
cat.dups <- cat.clean[duplicated(cat.clean$Info.1) | duplicated(cat.clean$Info.1, fromLast = TRUE),]  #locates any duplicated samples within cat.clean


#calculate standard deviation (error) using duplicates
#for anions
an.std <- matrix(NA, nrow = 1, ncol =7) #create empty matrix
for (i in 1:7){
  an.makeodd <- seq_len(nrow(an.dups)) %% 2                           #creating a sequence for the number of rows of an.dups
  an.diff <- abs(diff(as.numeric(an.dups[,i+4])))[an.makeodd == 1]    #taking the difference between duplicates
  an.diff <- na.omit(an.diff)                                         #omit NA's
  an.std[1,i] <- matrix(mean(an.diff))                                #insert error (mean of the dup diffs) into the matrix
}
an.std <- replace(an.std, an.std == "NaN", NA)
colnames(an.std) <- an.names

#for cations
cat.std <- matrix(NA, nrow = 1, ncol =6)
for (i in 1:6){
  cat.makeodd <- seq_len(nrow(cat.dups)) %% 2
  cat.diff <- abs(diff(as.numeric(cat.dups[,i+3])))[cat.makeodd == 1]
  cat.diff <- na.omit(cat.diff)
  cat.std[1,i] <- matrix(mean(cat.diff))
}
cat.std <- replace(cat.std, cat.std == "NaN", NA)
colnames(cat.std) <- cat.names


#create std columns
an.std.rep <- an.std[rep(1, nrow(an.conc)), ]        #repeats the an.std values vertically for the length of an.clean
cat.std.rep <- cat.std[rep(1, nrow(cat.conc)), ]        #repeats the cat.std values vertically for the length of cat.clean

#name them
colnames(an.std.rep) <- c("Fluoride STD","Chloride STD","Nitrite STD","Bromide STD","Nitrate STD","Phosphate STD","Sulfate STD")
colnames(cat.std.rep) <- c("Lithium STD","Sodium STD","Ammonium STD","Potassium STD","Magnesium STD","Calcium STD")

#detection limit workflow: 

#split the calibration data into concentration and area
# #for anions
an.cal.conc <- select(an_calibration, Anions.Fluoride.Concentration, Anions.Chloride.Concentration, Anions.Nitrite.as.N.Concentration, Anions.Bromide.Concentration,
                      Anions.Nitrate.as.N.Concentration, Anions.Phosphate.as.P.Concentration, Anions.Sulfate.Concentration)
an.cal.area <- select(an_calibration, Anions.Fluoride.Area, Anions.Chloride.Area, Anions.Nitrite.as.N.Area, Anions.Bromide.Area,
                      Anions.Nitrate.as.N.Area, Anions.Phosphate.as.P.Area, Anions.Sulfate.Area)

#for cations
cat.cal.conc <- select(cat_calibration, Cations.Lithium.Concentration, Cations.Sodium.Concentration, Cations.Ammonium.as.N.Concentration,
                       Cations.Potassium.Concentration, Cations.Magnesium.Concentration, Cations.Calcium.Area)
cat.cal.area <- select(cat_calibration, Cations.Lithium.Area, Cations.Sodium.Area, Cations.Ammonium.as.N.Area,
                       Cations.Potassium.Area, Cations.Magnesium.Area, Cations.Calcium.Area)


#create a linear model
#for anions
lm.an <- list()
for (i in 1:7){
  lm.an[[i]] <- lm(an.cal.area[,i] ~ an.cal.conc[,i])     #creates a linear model where area is the response/dependent variable,
}                                                   #and concentration is independent

#for cations
lm.cat <- list()
for (i in 1:6){
  lm.cat[[i]] <- lm(cat.cal.area[,i] ~ cat.cal.conc[,i])
}

#pull residuals and coefficients from the linear model
#for anions
sd.resid.an <- matrix(nrow = 1, ncol = length(lm.an)) #initiallizing empty matrix
slope.an <- matrix(nrow = 1, ncol = length(lm.an))    #initializing empty matrix
colnames(sd.resid.an) <- an.names
colnames(slope.an) <- an.names

for (t in 1:length(lm.an)){
  sd.resid.an[,t] <- sd(lm.an[[t]][["residuals"]])  #pulls residuals from the linear model
  slope.an[,t] <- lm.an[[t]][[1]][[2]]              #pulls slope coefficients from the linear model
}

#for cations
sd.resid.cat <- matrix(nrow = 1, ncol = length(lm.cat))
slope.cat <- matrix(nrow = 1, ncol = length(lm.cat))
colnames(sd.resid.cat) <- cat.names
colnames(slope.cat) <- cat.names

for (j in 1:6){
  sd.resid.cat[,j] <- sd(lm.cat[[j]][["residuals"]])
  slope.cat[,j] <- lm.cat[[j]][[1]][[2]]
}


#calculate detection limit
#for anions
DL.an <- matrix(nrow = 1, ncol = length(lm.an))  #initializing empty matrix
colnames(DL.an) <- an.names
for (k in 1:length(lm.an)){
  DL.an[,k] <- 3.3 * sd.resid.an[k]/slope.an[k]  #calculating DL using slope and sd of residuals
}

#for cations
DL.cat <- matrix(nrow = 1, ncol = length(lm.cat))
colnames(DL.cat) <- cat.names
for (k in 1:length(lm.cat)){
  DL.cat[,k] <- 3.3 * sd.resid.cat[k]/slope.cat[k]
}


#reprocess data
#for anions
for (y in 1:length(DL.an)){
  an.clean[[y+4]][an.clean[[y+4]] < DL.an[y]] <- -9999    #inserts -9999 if any samples are below DL
}

#for cations
for (x in 1:length(DL.cat))
  cat.clean[[x+4]][cat.clean[[x+4]] < DL.cat[x]] <- -9999    #inserts -9999 if any samples are below DL

#repeating DL columns for output
#for anions
DL.an.rep <- matrix(rep(DL.an, each = nrow(an.conc)), ncol = 7)
colnames(DL.an.rep) <- c("Fluoride DL","Chloride DL","Nitrite DL","Bromide DL","Nitrate DL","Phosphate DL","Sulfate DL")

#for cations
DL.cat.rep <- matrix(rep(DL.cat, each = nrow(cat.conc)), ncol = 6)
colnames(DL.cat.rep) <- c("Lithium DL","Sodium DL","Ammonium DL","Potassium DL","Magnesium DL","Calcium DL")


## Rename columns so that they can be sorted 
#full outputs
#anions
anion_processed_output <- cbind(an.clean, an.std.rep, DL.an.rep)
an.sorted.colnames <- sort(names(anion_processed_output))
anion_processed_output <- anion_processed_output[,an.sorted.colnames]
columns.to.move <- c("Ident","Info.1","Determination.start")
anion_processed_output <- anion_processed_output[, c(columns.to.move, setdiff(names(anion_processed_output), columns.to.move))]
anion_processed_output <- anion_processed_output %>% select(-Sample.type) # removing column

#cations
cation_processed_output <- cbind(cat.clean, cat.std.rep, DL.cat.rep)
cat.sorted.colnames <- sort(names(cation_processed_output))
cation_processed_output <- cation_processed_output[,cat.sorted.colnames]
columns.to.move <- c("Ident","Info.1","Determination.start")
cation_processed_output <- cation_processed_output[, c(columns.to.move, setdiff(names(cation_processed_output), columns.to.move))]
cation_processed_output <- cation_processed_output %>% select(-Sample.type) # removing column

# Export to CSV
output_folder <- file.path(getwd(), 'Processed_IC_Data')
#create folder if does not already exist
if (!dir.exists(output_folder)) {
  dir.create(output_folder)
}

# create export file name using file name provided at the beginning of the scipt
anion_file_processed <- sub("\\.csv$", "_processed.csv", anion_file)
cation_file_processed <- sub("\\.csv$", "_processed.csv", cation_file)

# export data using file name provided at beginning of script
write.csv(anion_processed_output, file = file.path(output_folder, anion_file_processed), row.names = FALSE)
write.csv(cation_processed_output, file = file.path(output_folder, cation_file_processed), row.names = FALSE)

```

#### No 'Ident' column (exported pre 15Oct25)

```{r}

#specify column names
an.names <- c("Fluoride","Chloride","Nitrite", "Bromide","Nitrate","Phosphate","Sulfate")
cat.names <- c("Lithium","Sodium","Ammonium","Potassium","Magnesium","Calcium")

##clean the data to remove standards, blanks, and invalid values
an.clean <- anions[which(anions$Sample.type == "Sample"),]    #checks the sample.type column and pulls only "Sample" values
cat.clean <- cations[which(cations$Sample.type == "Sample"),]

an.clean <- replace(an.clean, an.clean == 'invalid', NA)   #replaces any 'invalid' values with NA
cat.clean <- replace(cat.clean, cat.clean == 'invalid', NA)

##Doing the same cleaning but for the calibration run
an_calibration  <- an_calibration  %>% filter(grepl("Standard", `Sample.type`, ignore.case = TRUE))
cat_calibration <- cat_calibration %>% filter(grepl("Standard", `Sample.type`, ignore.case = TRUE))

an_calibration <- an_calibration[an_calibration$`Sample.type` != "Check standard 1", ]
cat_calibration <- cat_calibration[cat_calibration$`Sample.type` != "Check standard 1", ]

an_calibration <- replace(an_calibration, an_calibration == 'invalid', NA)   #replaces any 'invalid' values with NA
cat_calibration <- replace(cat_calibration, cat_calibration == 'invalid', NA)


#remove '-duplicate' from any sample names
an.clean$Info.1 <- gsub("-duplicate", "", as.character(an.clean$Info.1))
an.clean$Info.1 <- gsub("-Duplicate", "", as.character(an.clean$Info.1))
an.clean$Info.1 <- gsub("-DUPLICATE", "", as.character(an.clean$Info.1))
an.clean$Info.1 <- gsub(" DUPLICATE", "", as.character(an.clean$Info.1))
an.clean$Info.1 <- gsub(" Duplicate", "", as.character(an.clean$Info.1))
an.clean$Info.1 <- gsub(" duplicate", "", as.character(an.clean$Info.1))
an.clean$Info.1 <- gsub("_Duplicate", "", as.character(an.clean$Info.1))
an.clean$Info.1 <- gsub("_duplicate", "", as.character(an.clean$Info.1))
colnames(an.clean) <- c("Determination.start","Sample.type","Info.1","Fluoride Concentration","Chloride Concentration","Nitrite Concentration", "Bromide Concentration",
                       "Nitrate Concentration","Phosphate Concentration","Sulfate Concentration")

cat.clean$Info.1 <- gsub("-duplicate", "", as.character(cat.clean$Info.1))
cat.clean$Info.1 <- gsub("-Duplicate", "", as.character(cat.clean$Info.1))
cat.clean$Info.1 <- gsub("-DUPLICATE", "", as.character(cat.clean$Info.1))
cat.clean$Info.1 <- gsub(" duplicate", "", as.character(cat.clean$Info.1))
cat.clean$Info.1 <- gsub(" Duplicate", "", as.character(cat.clean$Info.1))
cat.clean$Info.1 <- gsub(" DUPLICATE", "", as.character(cat.clean$Info.1))
cat.clean$Info.1 <- gsub("_Duplicate", "", as.character(cat.clean$Info.1))
cat.clean$Info.1 <- gsub("_duplicate", "", as.character(cat.clean$Info.1))
cat.clean$Info.1 <- gsub(" Deuplicate", "", as.character(cat.clean$Info.1))
colnames(cat.clean) <- c("Determination.start","Sample.type","Info.1","Lithium Concentration","Sodium Concentration","Ammonium Concentration",
                        "Potassium Concentration","Magnesium Concentration","Calcium Concentration")

#create cation and anion concentration variables
an.conc <- subset(an.clean, select = -Sample.type)
cat.conc <- subset(cat.clean, select = -Sample.type)



#standard deviation workflow:

#find duplicates
an.dups <- an.clean[duplicated(an.clean$Info.1) | duplicated(an.clean$Info.1, fromLast = TRUE),]  #locates any duplicated samples within an.clean
cat.dups <- cat.clean[duplicated(cat.clean$Info.1) | duplicated(cat.clean$Info.1, fromLast = TRUE),]  #locates any duplicated samples within cat.clean


#calculate standard deviation (error) using duplicates
#for anions
an.std <- matrix(NA, nrow = 1, ncol =7) #create empty matrix
for (i in 1:7){
  an.makeodd <- seq_len(nrow(an.dups)) %% 2                           #creating a sequence for the number of rows of an.dups
  an.diff <- abs(diff(as.numeric(an.dups[,i+3])))[an.makeodd == 1]    #taking the difference between duplicates
  an.diff <- na.omit(an.diff)                                         #omit NA's
  an.std[1,i] <- matrix(mean(an.diff))                                #insert error (mean of the dup diffs) into the matrix
}
an.std <- replace(an.std, an.std == "NaN", NA)
colnames(an.std) <- an.names

#for cations
cat.std <- matrix(NA, nrow = 1, ncol =6)
for (i in 1:6){
  cat.makeodd <- seq_len(nrow(cat.dups)) %% 2
  cat.diff <- abs(diff(as.numeric(cat.dups[,i+3])))[cat.makeodd == 1]
  cat.diff <- na.omit(cat.diff)
  cat.std[1,i] <- matrix(mean(cat.diff))
}
cat.std <- replace(cat.std, cat.std == "NaN", NA)
colnames(cat.std) <- cat.names


#create std columns
an.std.rep <- an.std[rep(1, nrow(an.conc)), ]        #repeats the an.std values vertically for the length of an.clean
cat.std.rep <- cat.std[rep(1, nrow(cat.conc)), ]        #repeats the cat.std values vertically for the length of cat.clean

#name them
colnames(an.std.rep) <- c("Fluoride STD","Chloride STD","Nitrite STD","Bromide STD","Nitrate STD","Phosphate STD","Sulfate STD")
colnames(cat.std.rep) <- c("Lithium STD","Sodium STD","Ammonium STD","Potassium STD","Magnesium STD","Calcium STD")

#detection limit workflow: 

#split the calibration data into concentration and area
# #for anions
an.cal.conc <- select(an_calibration, Anions.Fluoride.Concentration, Anions.Chloride.Concentration, Anions.Nitrite.as.N.Concentration, Anions.Bromide.Concentration,
                      Anions.Nitrate.as.N.Concentration, Anions.Phosphate.as.P.Concentration, Anions.Sulfate.Concentration)
an.cal.area <- select(an_calibration, Anions.Fluoride.Area, Anions.Chloride.Area, Anions.Nitrite.as.N.Area, Anions.Bromide.Area,
                      Anions.Nitrate.as.N.Area, Anions.Phosphate.as.P.Area, Anions.Sulfate.Area)

#for cations
cat.cal.conc <- select(cat_calibration, Cations.Lithium.Concentration, Cations.Sodium.Concentration, Cations.Ammonium.as.N.Concentration,
                       Cations.Potassium.Concentration, Cations.Magnesium.Concentration, Cations.Calcium.Area)
cat.cal.area <- select(cat_calibration, Cations.Lithium.Area, Cations.Sodium.Area, Cations.Ammonium.as.N.Area,
                       Cations.Potassium.Area, Cations.Magnesium.Area, Cations.Calcium.Area)


#create a linear model
#for anions
lm.an <- list()
for (i in 1:7){
  lm.an[[i]] <- lm(an.cal.area[,i] ~ an.cal.conc[,i])     #creates a linear model where area is the response/dependent variable,
}                                                   #and concentration is independent

#for cations
lm.cat <- list()
for (i in 1:6){
  lm.cat[[i]] <- lm(cat.cal.area[,i] ~ cat.cal.conc[,i])
}

#pull residuals and coefficients from the linear model
#for anions
sd.resid.an <- matrix(nrow = 1, ncol = length(lm.an)) #initiallizing empty matrix
slope.an <- matrix(nrow = 1, ncol = length(lm.an))    #initializing empty matrix
colnames(sd.resid.an) <- an.names
colnames(slope.an) <- an.names

for (t in 1:length(lm.an)){
  sd.resid.an[,t] <- sd(lm.an[[t]][["residuals"]])  #pulls residuals from the linear model
  slope.an[,t] <- lm.an[[t]][[1]][[2]]              #pulls slope coefficients from the linear model
}

#for cations
sd.resid.cat <- matrix(nrow = 1, ncol = length(lm.cat))
slope.cat <- matrix(nrow = 1, ncol = length(lm.cat))
colnames(sd.resid.cat) <- cat.names
colnames(slope.cat) <- cat.names

for (j in 1:6){
  sd.resid.cat[,j] <- sd(lm.cat[[j]][["residuals"]])
  slope.cat[,j] <- lm.cat[[j]][[1]][[2]]
}


#calculate detection limit
#for anions
DL.an <- matrix(nrow = 1, ncol = length(lm.an))  #initializing empty matrix
colnames(DL.an) <- an.names
for (k in 1:length(lm.an)){
  DL.an[,k] <- 3.3 * sd.resid.an[k]/slope.an[k]  #calculating DL using slope and sd of residuals
}

#for cations
DL.cat <- matrix(nrow = 1, ncol = length(lm.cat))
colnames(DL.cat) <- cat.names
for (k in 1:length(lm.cat)){
  DL.cat[,k] <- 3.3 * sd.resid.cat[k]/slope.cat[k]
}


#reprocess data
#for anions
for (y in 1:length(DL.an)){
  an.clean[[y+3]][an.clean[[y+3]] < DL.an[y]] <- -9999    #inserts -9999 if any samples are below DL
}

#for cations
for (x in 1:length(DL.cat))
  cat.clean[[x+3]][cat.clean[[x+3]] < DL.cat[x]] <- -9999    #inserts -9999 if any samples are below DL


#repeating DL columns for output
#for anions
DL.an.rep <- matrix(rep(DL.an, each = nrow(an.conc)), ncol = 7)
colnames(DL.an.rep) <- c("Fluoride DL","Chloride DL","Nitrite DL","Bromide DL","Nitrate DL","Phosphate DL","Sulfate DL")

#for cations
DL.cat.rep <- matrix(rep(DL.cat, each = nrow(cat.conc)), ncol = 6)
colnames(DL.cat.rep) <- c("Lithium DL","Sodium DL","Ammonium DL","Potassium DL","Magnesium DL","Calcium DL")


## Rename columns so that they can be sorted 
#full outputs
#anions
anion_processed_output <- cbind(an.clean, an.std.rep, DL.an.rep)
an.sorted.colnames <- sort(names(anion_processed_output))
anion_processed_output <- anion_processed_output[,an.sorted.colnames]
columns.to.move <- c("Info.1","Determination.start")
anion_processed_output <- anion_processed_output[, c(columns.to.move, setdiff(names(anion_processed_output), columns.to.move))]
anion_processed_output <- anion_processed_output %>% select(-Sample.type) # removing column


#cations
cation_processed_output <- cbind(cat.clean, cat.std.rep, DL.cat.rep)
cat.sorted.colnames <- sort(names(cation_processed_output))
cation_processed_output <- cation_processed_output[,cat.sorted.colnames]
columns.to.move <- c("Info.1","Determination.start")
cation_processed_output <- cation_processed_output[, c(columns.to.move, setdiff(names(cation_processed_output), columns.to.move))]
cation_processed_output <- cation_processed_output %>% select(-Sample.type) # removing column

# Export to CSV
output_folder <- file.path(getwd(), 'Processed_IC_Data')
#create folder if does not already exist
if (!dir.exists(output_folder)) {
  dir.create(output_folder)
}

# create export file name using file name provided at the beginning of the scipt
anion_file_processed <- sub("\\.csv$", "_processed.csv", anion_file)
cation_file_processed <- sub("\\.csv$", "_processed.csv", cation_file)

# export data 
write.csv(anion_processed_output, file = file.path(output_folder, anion_file_processed), row.names = FALSE)
write.csv(cation_processed_output, file = file.path(output_folder, cation_file_processed), row.names = FALSE)

```

#### Run this cell if you need to manually export the final files.

```{r}
write.csv(anion_processed_output,"/Users/isabellawelk/Documents/GitHub/Detection-Limits/ICRunData/Outputs/26July24_Anion.csv", row.names = FALSE)

write.csv(cation_processed_output,"/Users/isabellawelk/Documents/GitHub/Detection-Limits/ICRunData/Outputs/31July24_Cation.csv", row.names = FALSE)

```
